---
title: "Data_analysis"
output: html_document
date: "2025-04-17"
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
install.packages("dunn.test")
library(dunn.test)
```

```{r}
NHANES_analysis <- read.csv("/Users/renxichen/Documents/ANTH241 Project/clean_data")
NHANES_analysis
```
factor
```{r}
factor(c(NHANES_analysis$Little_interest_do_things, NHANES_analysis$Controlling_or_losing_weight, NHANES_analysis$Difficulty_concentrating))
factor(c(NHANES_analysis$Little_interest_frequency, NHANES_analysis$Drug_use_frequency, NHANES_analysis$Marijuana_use, NHANES_analysis$Sleep_timeEL), order=TRUE)
head(NHANES_analysis)

```

frequency tables
```{r}
library(questionr)
freq(NHANES_analysis$Marijuana_use, cum = TRUE, na.omit(NHANES_analysis$Marijuana_use))
freq(NHANES_analysis$Sleep_timeEL, cum = TRUE)
freq(NHANES_analysis$Drug_use_frequency, cum = TRUE)
freq(NHANES_analysis$Little_interest_frequency, cum = TRUE)
```


mean, sd, count for sedentary activity (already finished drop_na and filter in data cleaning)
```{r}
NHANES_analysis%>%
  summarize(mean(Sedentary_activity), sd(Sedentary_activity), count=n())
```

generate mean, sd, and count by little interest frequency
```{r}
NHANES_analysis%>%
  group_by(Little_interest_frequency)%>%
  summarize(mean(Sedentary_activity), sd(Sedentary_activity), count=n())
```


boxplot of sedentary activity by little interest frequency
```{r}
ggplot(NHANES_analysis, aes(x=Little_interest_frequency, y=Sedentary_activity))+geom_boxplot(fill="orange")+labs(x="Little Interest Frequency Levels", y="Sedentary Activity Minutes", title="Sedentary Minutes by Little Interest Frequency")
```
This is the graph showing how many minutes people in different interest levels do sedentary activity. 

Kruskal Wallis and Dunn test
```{r}
dunn.test(x=NHANES_analysis$Sedentary_activity, g=NHANES_analysis$Little_interest_frequency, kw=TRUE)
```
There is a significant different in group of nearly everyday (have little interest in doing things) and group of not at all(p = 0.03), group of not at all and several days (p = 0.03).

f test and t test for difficulty concentrating
```{r}
var.test(Little_interest_do_things~Difficulty_concentrating, data=NHANES_analysis, alternative="two.sided")
t.test(Little_interest_do_things~Difficulty_concentrating, var.equal=TRUE, data=NHANES_analysis)
```
p = 0.9, there is not significant difference in having difficulty concentrating or not between groups of people who have different levels of interest doing things.

f test and t test for controlling or losing weight
```{r}
var.test(Little_interest_do_things~Controlling_or_losing_weight, data=NHANES_analysis, alternative="two.sided")
t.test(Little_interest_do_things~Controlling_or_losing_weight, var.equal=TRUE, data=NHANES_analysis)
```
p = 0.27, there is not significant difference in controlling or losing weight between groups of people who have different levels of interest doing things.

chi square of drug use and little interest do things
```{r}
Drug_Little_table <- xtabs(~ Drug_use_frequency + Little_interest_frequency, data=NHANES_analysis)
Drug_Little_table
prop.table(Drug_Little_table)*100
chisq.test(Drug_Little_table)
```
p = 0.15, there is not significant difference in drug using among groups of people have different levels of interest doing things.

little interest in doing things
```{r}
ggplot(NHANES_analysis, aes(x=Little_interest_frequency))+geom_bar(fill="pink")+labs(title="Distribution of Having Little Interest in Doing Things", x="Levels", y="Counts")+geom_text(stat = 'count',aes(label=..count..),vjust=-0.5)
```

sedentary activity
```{r}
ggplot(NHANES_analysis, aes(x=Sedentary_activity))+geom_bar(fill="black")+labs(title = "Distribution of Sedentary Activity", x="Sedentary Activity", y="Counts")
```
marijuana use
```{r}
ggplot(NHANES_analysis, aes(x=Drug_use_frequency))+geom_bar(fill="pink")+labs(title = "Drug Use Frequency", x="Drug Use Frequency", y="Counts") + geom_text(stat = 'count',aes(label=..count..),vjust=-0.5)
```
Difficulty concentrating
```{r}
ggplot(NHANES_analysis, aes(x=Difficulty_concentrating))+geom_bar(fill="pink")+labs(title = "Difficulty Concentrating", x="Difficulty Concentrating", y="Counts") + geom_text(stat = 'count',aes(label=..count..),vjust=-0.5)
```

Controlling or Losing Weight
```{r}
ggplot(NHANES_analysis, aes(x=Controlling_or_losing_weight))+geom_bar(fill="pink")+labs(title = "Controlling or Losing Weight", x="Controlling or Losing Weight", y="Counts") + geom_text(stat = 'count',aes(label=..count..),vjust=-0.5)
```


Distribution of Sleep early or late
```{r}
ggplot(NHANES_analysis, aes(x=Little_interest_frequency, fill=Sleep_timeEL))+geom_bar(position="dodge")+labs(title="Distribution of Sleep Early or Late in Levels", x="Levels of Having Little Interest in Doing Things", y="Sleep Time")
```
This is a bar chart showing the distribution of different sleep time in each levels of having little interest in doing things. 

```{r}
NHANES_analysis%>%
  group_by(Sleep_timeEL)%>%
  summarize(count=n())
```

little interest and sedentary
```{r}
ggplot(aes(x=Little_interest_do_things, y=Sedentary_activity), data=NHANES_analysis) + geom_point(color="orange") + geom_smooth(method=lm, color="black") + labs(x="Little Interest Levels", y="Sedentary Activity", title="Little Interest and Sedentary Minutes") + theme_bw()
cor.test(NHANES_analysis$Little_interest_do_things, NHANES_analysis$Sedentary_activity, method="pearson")
```
p = 0.007, there is a significant difference among sedentary activity minutes in different levels of having interest in doing things. In this graph, the levels are: 0 = not at all, 1 = several days, 2 = more than half of the days, 3 = nearly everyday.

little interest and marijuana use
```{r}
ggplot(aes(x=Little_interest_do_things, y=Marijuana_use), data=NHANES_analysis) + geom_point(color="orange") + geom_smooth(method=lm, color="black") + labs(x="Little Interest Levels", y="Marijuana Use Levels", title="Little Interest and Drug Use") + theme_bw()
cor.test(NHANES_analysis$Little_interest_do_things, NHANES_analysis$Marijuana_use, method="pearson")
```
p = 0.15, there is not significant difference among marijuana using in different levels of having interest in doing things. In this graph, the levels of little interest are: 0 = not at all, 1 = several days, 2 = more than half of the days, 3 = nearly everyday, the levels of marijuana use are: 1 = rarely, 2 = sometimes, 3 = often, 4 = usually, 5 = always. 


linear regression model of little interest and sedentary activity
```{r}
Little_Sedentary_fit <- lm(Little_interest_do_things~Sedentary_activity, data=NHANES_analysis)
summary(Little_Sedentary_fit)
```
y(little interest) = 0.33 + sedentary activity (0.0004)
r squared value: 0.012
p value = 0.007 which is significant

odds ratio of little interest and usual sleep time
```{r}
Sleep_Little_odds_table <- table(NHANES_analysis$Sleep_timeEL, NHANES_analysis$Little_interest_frequency)
oddsratio(Sleep_Little_odds_table)
ggplot(data=NHANES_analysis) + geom_mosaic(aes(x=product(Sleep_timeEL), fill=Little_interest_frequency), na.rm=TRUE, show.legend=FALSE) + labs(title="Sleep Time by Little Interest Levels", x="Sleep Time", y="Little Interest Levels")
```
compare to early, odds ratios are: irregular: 3.35, late: 1.14, normal/medium: 2.21, super late: 1.39.
there is no p value < 0.05 which means that there is not significant difference among any of them. 

odds ratio of little interest and drug use
```{r}
Drug_Little_odds_table <- table(NHANES_analysis$Drug_use_frequency, NHANES_analysis$Little_interest_frequency)
oddsratio(Drug_Little_odds_table)
ggplot(data=NHANES_analysis) + geom_mosaic(aes(x=product(Drug_use_frequency), fill=Little_interest_frequency), na.rm=TRUE, show.legend=FALSE) + labs(title="Drug Use by Little Interest Levels", x="Drug Use Frequency", y="Little Interest Levels")
```
compare to always, odds ratios are: often: 0.54, rarely: 0.84, sometimes: 1.55, usually: 1.19
there is no p value < 0.05 which means that there is not significant difference among any of them

odds ratio of little interest and difficulty concentrating
```{r}
Concentrate_Little_odds_table <- table(NHANES_analysis$Little_interest_frequency, NHANES_analysis$Difficulty_concentrating)
oddsratio(Concentrate_Little_odds_table)
ggplot(data=NHANES_analysis) + geom_mosaic(aes(x=product(Little_interest_frequency), fill=Difficulty_concentrating), na.rm=TRUE, show.legend=FALSE) + labs(title="Difficulty Concentrating by Little Interest Levels", x="Little Interest Levels", y="Difficulty Concentrating")
```
compare to more than half of the days, odds ratios are: nearly everyday: 0.35, not at all: 0.92, several days: 1.43.
there is no p value < 0.05 which means there is not significant difference among any of them.

odds ratio of little interest and controlling or losing weight
```{r}
Weight_Little_odds_table <- table(NHANES_analysis$Little_interest_frequency, NHANES_analysis$Controlling_or_losing_weight)
oddsratio(Weight_Little_odds_table)
ggplot(data=NHANES_analysis) + geom_mosaic(aes(x=product(Little_interest_frequency), fill=Controlling_or_losing_weight), na.rm=TRUE, show.legend=FALSE) + labs(title="Controlling or losing weight by Little Interest Levels", x="Little Interest Levels", y="Controlling or Losing Weight")
```
compare to more than half of the days, odds ratios are: nearly everyday: 1.10,not at all: 1.39, several days: 1.22.
there is no p value < 0.05 which means there is not significant difference among any of them. 
